---
title: "README"
author: "Freddie Mckendrick"
date: "10/02/2021"
output: html_document
---

I have tried to contain any information about what is in each file within this R project here 

The objective of this project is to quantify the repeatability of provisioning rate in the Seychelles warbler
All files and analysis are saved in the GitHub repository
https://github.com/FreddieMckendrick/ProvisioningRateRepeatability.git
Folder names and contents are outlined below 

Access_queries 
* Word ducements containing SQL code for the extraction of the required data from the Seychelles warbler data base
* Documents give both the SQL code and instructions for how the data should be extracted and what queries should be saved as 

Data
* Contains five different datasets
* These include the provisioning data, two files for Not seen but alive individuals (NSA) and two files used to calculate territory quality extracted from the database by Ellie Chesterton
* These data sets are imported into the script as tibbles 

Figures
* Contains PDF and Image outputs of all figures to be used in thesis 
* Repeatability estimates
* Parameter estimates for fixed effects within Model

Outputs 
* Includes any saved model outputs and processed data sets that were filtered and used within the analysis 

Reports
* Includes R Markdown files 


The versions of each package used in the analysis are outlined below 

```{r}
library(tidyverse)
library(Matrix)
library(coda)
library(ape)
library(MCMCglmm)
library(beepr)
library(rptR)
library(lme4)
library(insight)
library(parallel)
library(readxl)
library(lubridate)
library(magrittr)
```

```{r}
packageVersion("tidyverse")
packageVersion("Matrix")
packageVersion("coda")
packageVersion("ape")
packageVersion("MCMCglmm")
packageVersion("beepr")
packageVersion("rptR")
packageVersion("lme4")
packageVersion("insight")
packageVersion("parallel")
packageVersion("readxl")
packageVersion("lubridate")
packageVersion("magrittr")

```

